---
title: "HM1MuZepeng"
author: "Zepeng Mu, UCID: 12180655"
date: "9/27/2017"
output: html_document
---
##Lab 1 Homework
<span style="color:red"> The homework is due by 11:59 PM on the Sunday following your lab. Provide your answers in an Rmd file with the contents of this homework. Call the Rmd file “HW1LastnameFirstname.Rmd” and submit it on Canvas for your T.A. to grade. Remember to comment your code as we did for Lab 1. \span

<span style="color:green"> 1. Research built-in functions in R. Successfully execute 5 functions that were not explicitly covered in this lab. Through comments in your code, briefly describe what the function does. \span

<span style="color:blue">
```{r}
v1 = c(1:5)
#The function max() accepts a vector as its input and finds the maximum value in the vector.
max(v1)

#The function max() accepts a vector as its input and finds the minimum value in the vector.
min(v1)

#The function range() accepts a vector as its input and returns a vector consisting of the min and the max value in the input.
range(v1)

v2 = c(1, 3, 4, 6, 5, 2)
#The function sort() accepts a vector as its input and sorts the numbers in the vector either in an 'increasing' order (default) or a 'decreasing' order, along with other features such as the method used for sorting.
sorted = sort(v2)
sorted

#The function rnorm generates a vector of normally distributed random numbers of length n with a mean value and a sd value designated.
gauss = rnorm(n = 100, mean = 0, sd = 1)
gauss
```
\span

<span style="color:green"> 2. `While()` loops will iterate as long as the expression within parentheses evaluates to `TRUE`. Write a simple `while()` loop in R. What would happen if the code below were executed? Explain. Modify the code so that it executes a functional while loop (that only prints a finite number of lines).  \span
```{r}
#x <- 1  
#while(x <=  5){  
#	print(x + x)  
#}  
```

<span style="color:blue">
A simple `while()` loop would look like this. \span

```{r}
i = 1
while(i < 7) {
  print(i ^ 2)
  i = i + 1
}
```

<span style="color:blue">
The `while()` loop in the question simply would not stop and keep printing "2" since the value of x does not changed. To make it normal, it can be written in the following way: \span

```{r}
x <- 1  
while(x <=  5){  
	print(x + x)  
  x = x + 1
}  
```


<span style="color:green"> 3. (a) Using a for loop convert each pH from 1-14 into concentrations of H+. Store the list of [H+] in a vector. Remember that pH is equal to the –log10([H+]). To define an empty vector of length n and fill in values one at a time, see the following example.\span
```{r}
myVec <- numeric(4)
print(myVec)
for(index in seq_along(myVec)) {
    myVec[index] <- index^2
}
print(myVec)
```

<span style="color:blue">
My code is:
\span

```{r}
pH = numeric(length(1:14)) #stores [H+]

for(index in seq_along(pH)) {
  pH[index] = 10 ^ (-index)
}

print(pH)
```

<span style="color:green"> (b) Now write a function called H_from_pH that given a pH value, outputs the concentration of H+. Also, write an analogous function OH_from_pH to get [OH-] from the pH value.\span

<span style="color:blue">
My code is:
\span

```{r}
H_from_pH = function(pH) {
  return(10 ^ (-pH))
}

OH_from_pH = function(pH) {
  return(10 ^ (14 - pH))
}
```

<span style="color:green"> 4. Below is the formula for calculating standard deviation. Translate this formula into an R function that returns the standard deviation of a vector of numbers. Execute this function on a vector 1 through 100 with step size of 0.25 (remember
`seq()` !). Check that the results of your function match R’s built-in function `sd()`.  \span

\[s_x = \sqrt{\frac{\sum_{i=1}^n (x_i - \overline{x})^2}{n-1}}\]

<span style="color:green"> Where $n$ is the number of data points, $\overline{x}$ is the mean of the $x_i$, and $x_i$ each of the values of the data. \span  

<span style="color:blue">
My code is:
\span

```{r}
calcSD = function(Vec) {
  m = mean(Vec) #calculate mean
  
  sumOfVariance = 0
  for(element in Vec) {
    sumOfVariance = sumOfVariance + (element - m) ^ 2
  }
  
  SD = sqrt(sumOfVariance/(length(Vec) - 1))
  return(SD)
}

test = seq(1, 100, 0.25)

if(calcSD(test) == sd(test)) {
  print(paste0("sd = ", sd(test), ". The function works well."))
}

```

<span style="color:green"> 5. Write two functions `countG()` and `countC()` that, given a DNA sequence (like the example given below), output the number of guanines and the number of cytosines, respectively. Call your functions on `exampleSequence`.\span  

```{r}
exampleSequence = sample(c("a", "t", "c", "g"), 100, replace = TRUE)
exampleSequence

#My code starts from here:
countG = function(Seq) {
  return(length(Seq[Seq == "g"]))
}

countC = function(Seq) {
  return(length(Seq[Seq == "c"]))
}

countG(exampleSequence)
countC(exampleSequence)
```
